---
title: "terraZonal vs terraExtract vs exactExtractr (WDPA)"
author: "Om Prakash Bhandari"
date: "28/03/2022"
output: workflowr::wflow_html
---

## Introduction

There are several ways and techniques to process raster datasets - using different packages or the functionalities provided by those packages. Depending on the package and their functions, the time taken to process the rasters also varies accordingly. So, this study will focus on three important functions which are used widely to process rasters for three different types of polygons.

1. terra::zonal
2. terra::extract
3. exactextractr::exact_extract 

### Polygons - area of interest (aoi)

1. wdpa polygons (7495)
2. simplified wdpa polygons (4722)
3. fishnet gridded cells (216425)

Let's see how these three different aoi polygons look like:

```{r leafletdata, message=FALSE, warning=FALSE, include=FALSE}

library("sf")
library("terra")
library("leaflet")
library("exactextractr")
library("tidyverse")
library("tidyr")
library("ggplot2")
library("leaflet")
library("leaflet.extras")
library("leaflet.extras2")
library("ggsci")
library("scales")
library("htmltools")
library("RColorBrewer")
library("plotly")
library("rmarkdown")
library("magrittr")

# load WDPA polygon
pa.sf <- read_sf("../../datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-04-22_allPAs_valid.gpkg")

# load simplified WDPA polygon
spa.sf <- read_sf("../../datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-04-22_allPAs_valid_simplified.gpkg")

# load fishnet polygons
fishnet.sf <- read_sf("../../datalake/mapme.protectedareas/processing/fishnet/honeycomb_5_sqkm_kfw.gpkg")
fishnet.sf <- st_transform(fishnet.sf, 4326)


# subset data for only map
pa_map <- pa.sf %>%
  dplyr::filter(ISO3 %in% "DOM")
spa_map <- spa.sf %>%
  dplyr::filter(ISO3 %in% "DOM")
fish_map <- st_crop(fishnet.sf, pa_map)
# visualize
my_map <-
  leaflet() %>%
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addPolygons(data = pa_map,opacity = 0.9,color = "orange", group = "WDPA polygons",label = ~htmlEscape(WDPAID),weight = 1)%>%
  addPolygons(data = spa_map,opacity = 1,color = "green", group = "Simplified WDPA polygons",label = ~htmlEscape(WDPAID),weight = 1)%>%
  addPolygons(data = fish_map,opacity = 1,color = "blue", group = "Fishnets",label = ~htmlEscape(poly_id),weight = 1)%>%
  addFullscreenControl() %>%
  addLayersControl(
    baseGroups = c("CartoDB", "Satellite"),
    overlayGroups = c("WDPA polygons","Simplified WDPA polygons","Fishnets"),
    options = layersControlOptions(collapsed = FALSE))

```

```{r map, message=FALSE, warning=FALSE, include=TRUE}

my_map
```

### Raster properties

The raster used in this study is worldpop global population raster from year 2020. 

Lets see the properties of raster:

```{r rasload, message=FALSE, warning=FALSE, include=FALSE}

population_raster <- rast("../../datalake/mapme.protectedareas/input/world_pop/global_mosaic2020.tif")
```

```{r ras, message=FALSE, warning=FALSE, include=TRUE}

population_raster
```


### Processing time to compute raster statistics

After pocessing the population raster layer for three different study areas using three above-mentioned functions, we got the following result (here in the graphical form) considering the elapsed time.

```{r plot, message=FALSE, warning=FALSE, include=FALSE}

df.result <- read.csv("data/results/01_zon_v_textr_v_extr.csv")
df.result.long <- pivot_longer(df.result,
                               cols = c("WDPA_7495", "Simplified_WDPA_4722", "Fishnet_216425"))
result.plot <- ggplot(data = df.result.long, aes(x=name, y=value, fill=Functions)) + 
  geom_col(position = 'dodge') +
  xlab('Type of Polygons') +
  ylab('Time elapsed (s)') +
  ggtitle('comparison of time taken by three different functions to compute raster statistics')
```

```{r plot2, message=FALSE, warning=FALSE, include = TRUE, fig.width = 12}

result.plot
```
