---
title: "sequential vs parallelization (WDPA)"
author: "Om Prakash Bhandari"
date: "5/18/2022"
output: workflowr::wflow_html
---

## Introduction

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}

library("sf")
library("terra")
library("tidyr")
library("ggplot2")
library("tidyverse")
library("leaflet")
library("exactextractr")
library("leaflet.extras")
library("leaflet.extras2")
library("ggsci")
library("scales")
library("htmltools")
library("RColorBrewer")
library("plotly")
library("rmarkdown")
library("magrittr")
```

In this study, we are going to check how the processing time will differ - (1) sequential approach (2) parallel approach while processing rasters with varying resolutions and for different study areas:

1. wdpa polygons (167)
2. simplified wdpa polygons (152)
3. fishnet gridded cells (14338)

Let's see how these three different aoi polygons look like:

```{r leafletdata, message=FALSE, warning=FALSE, include=FALSE}

# load WDPA polygon
pa.sf <- read_sf("../sample_wdpa.gpkg")

# load simplified WDPA polygon
spa.sf <- read_sf("../sample_simp-wdpa.gpkg")

# load fishnet polygons
fishnet.sf <- read_sf("../sample_wdpa-fishnets.gpkg")

# visualize
my_map <-
  leaflet() %>%
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addPolygons(data = pa.sf,opacity = 0.9,color = "orange", group = "WDPA polygons",label = ~htmlEscape(WDPAID),weight = 1)%>%
  addPolygons(data = spa.sf,opacity = 1,color = "green", group = "Simplified WDPA polygons",label = ~htmlEscape(WDPAID),weight = 1)%>%
  addPolygons(data = fishnet.sf,opacity = 1,color = "blue", group = "Fishnets",label = ~htmlEscape(poly_id),weight = 1)%>%
  addFullscreenControl() %>%
  addLayersControl(
    baseGroups = c("CartoDB", "Satellite"),
    overlayGroups = c("WDPA polygons","Simplified WDPA polygons","Fishnets"),
    options = layersControlOptions(collapsed = FALSE))

```

```{r map, message=FALSE, warning=FALSE, include=TRUE}

my_map
```

Four raster objects with varying resolutions were generated across the Bolivia region to carry out this study.
The details of the rasters thus created follows:

a) 0.1 degree (~11.1 km) = size 71.7 KB
b) 0.01 degree (~1.11 km) = size 6.9 MB
c) 0.001 degree (~111 m) = size 691.9 MB
d) 0.0005 degree (~55.5 m) = size 2.7 GB


## Results

To carry out this analysis, all the results are aggregated and group by raster resolutions. Thus, for specific raster resolution, we could recommend what could be the most suitable combination.

### 0.1 degree resolution (~11.1 km)

```{r 0.1 degree final, message=FALSE, warning=FALSE, include=FALSE}

final01 <- read.csv("data/results/03_seq_v_par_wdpa/final_01.csv")
long <- tidyr::pivot_longer(final01,
                             cols = c("WDPA_167", "Simplified_WDPA_152", "Fishnets_14338"))

result <- ggplot(data = long, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values=c("deepskyblue", "deepskyblue1", "deepskyblue2", "deepskyblue3", "deepskyblue4",
                             "darkgoldenrod1", "darkgoldenrod2", "darkgoldenrod3", "darkgoldenrod", "darkgoldenrod4", "chartreuse", "chartreuse1", "chartreuse2", "chartreuse3", "chartreuse4")) +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)") +
  ggtitle("time taken by three different functions and cores to compute raster statistics (0.1 degree)")

# w/o sequential
f <- final01[ ,-4]
longf <- tidyr::pivot_longer(f,
                             cols = c("WDPA_167", "Simplified_WDPA_152"))
result_new <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values=c("deepskyblue", "deepskyblue1", "deepskyblue2", "deepskyblue3", "deepskyblue4",
                             "darkgoldenrod1", "darkgoldenrod2", "darkgoldenrod3", "darkgoldenrod", "darkgoldenrod4", "chartreuse", "chartreuse1", "chartreuse2", "chartreuse3", "chartreuse4")) +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)") +
  ggtitle("time taken by three different functions and cores to compute raster statistics (0.1 degree)")
```

```{r 0.1 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result
```

```{r 0.1 degree final plot 2, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result_new
```

### 0.01 degree resolution (~1.11 km)

```{r 0.01 degree final, message=FALSE, warning=FALSE, include=FALSE}

final001 <- read.csv("data/results/03_seq_v_par_wdpa/final_001.csv")
long <- tidyr::pivot_longer(final001,
                             cols = c("WDPA_167", "Simplified_WDPA_152", "Fishnets_14338"))

result <- ggplot(data = long, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values=c("deepskyblue", "deepskyblue1", "deepskyblue2", "deepskyblue3", "deepskyblue4",
                             "darkgoldenrod1", "darkgoldenrod2", "darkgoldenrod3", "darkgoldenrod", "darkgoldenrod4", "chartreuse", "chartreuse1", "chartreuse2", "chartreuse3", "chartreuse4")) +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)") +
  ggtitle("time taken by three different functions and cores to compute raster statistics (0.01 degree)")

# w/o terra zonal (4,8,16)
f <- final001[ ,-4]
longf <- tidyr::pivot_longer(f,
                             cols = c("WDPA_167", "Simplified_WDPA_152"))

result_new <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values=c("deepskyblue", "deepskyblue1", "deepskyblue2", "deepskyblue3", "deepskyblue4",
                             "darkgoldenrod1", "darkgoldenrod2", "darkgoldenrod3", "darkgoldenrod", "darkgoldenrod4", "chartreuse", "chartreuse1", "chartreuse2", "chartreuse3", "chartreuse4")) +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)") +
  ggtitle("time taken by three different functions and cores to compute raster statistics (0.01 degree)")
```

```{r 0.01 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result
```

_terra zonal(4,8,16) have created much spikes over the graphs, so lets remove zonal and plot only terra extract and exact extract._

```{r 0.01 degree final plot 2, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result_new
```


### 0.001 degree resolution (~111 m)

```{r 0.001 degree final, message=FALSE, warning=FALSE, include=FALSE}

final0001 <- read.csv("data/results/03_seq_v_par_wdpa/final_0001.csv")
long <- tidyr::pivot_longer(final0001,
                             cols = c("WDPA_167", "Simplified_WDPA_152", "Fishnets_14338"))

result <- ggplot(data = long, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values=c("deepskyblue", "deepskyblue1", "deepskyblue2", "deepskyblue3", "deepskyblue4",
                             "darkgoldenrod1", "darkgoldenrod2", "darkgoldenrod3", "darkgoldenrod", "darkgoldenrod4")) +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)") +
  ggtitle("time taken by three different functions and cores to compute raster statistics (0.001 degree)")
```

```{r 0.001 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result
```


### 0.0005 degree resolution (~55.5 m)

```{r 0.0005 degree final, message=FALSE, warning=FALSE, include=FALSE}

final00005 <- read.csv("data/results/03_seq_v_par_wdpa/final_00005.csv")
long <- tidyr::pivot_longer(final00005,
                             cols = c("WDPA_167", "Simplified_WDPA_152", "Fishnets_14338"))

result <- ggplot(data = long, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values=c("deepskyblue", "deepskyblue1", "deepskyblue2", "deepskyblue3", "deepskyblue4",
                             "darkgoldenrod1", "darkgoldenrod2", "darkgoldenrod3", "darkgoldenrod", "darkgoldenrod4")) +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)") +
  ggtitle("time taken by three different functions and cores to compute raster statistics (0.0005 degree)")
```

```{r 0.0005 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result
```



## Conclusion

|**Rasters**|**size**|**Combination of functions and methods**|
|-----------|--------|----------------------------------------|
|0.1 degree (~11.1 km) |71.7 KB|1. terra zonal in single asset mode|
|0.01 degree (~1.11 km) |6.9 MB|1. terra zonal in single asset mode </br> 2. exact extract in single asset mode|
|0.001 degree (~111 m)|691.9 MB|1. for wdpa polygons, exact extract with 16 cores </br> 2. for fishnets, exact extract single asset|
|0.0005 degree (~55.5 m)|2.7 GB|1. for wdpa polygons, exact extract with 16 cores </br> 2. for fishnets, exact extract single asset|

**terra zonal did not process very small polygons in many occasions, so if your study area contains very small areas, it is recommended not to opt terra zonal.**

