---
title: "sequential vs parallelization (WDPA)"
author: "Om Prakash Bhandari"
date: "5/18/2022"
output: workflowr::wflow_html
---

## Introduction

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}

library("sf")
library("terra")
library("tidyr")
library("ggplot2")
library("tidyverse")
library("leaflet")
library("exactextractr")
library("leaflet.extras")
library("leaflet.extras2")
library("ggsci")
library("scales")
library("htmltools")
library("RColorBrewer")
library("plotly")
library("rmarkdown")
library("magrittr")
```

In this study, we are going to check how the processing time will differ - (1) sequential approach (2) parallel approach while processing rasters with varying resolutions and for different study areas:

1. wdpa polygons (167)
2. simplified wdpa polygons (167)
3. fishnet gridded cells (14338)

Let's see how these three different aoi polygons look like:

```{r leafletdata, message=FALSE, warning=FALSE, include=FALSE}

# load WDPA polygon
pa.sf <- read_sf("../sample_wdpa.gpkg")

# load simplified WDPA polygon
spa.sf <- read_sf("../sample_simp-wdpa.gpkg")

# load fishnet polygons
fishnet.sf <- read_sf("../sample_wdpa-fishnets.gpkg")

# visualize
my_map <-
  leaflet() %>%
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addPolygons(data = pa.sf,opacity = 0.9,color = "orange", group = "WDPA polygons",label = ~htmlEscape(WDPAID),weight = 1)%>%
  addPolygons(data = spa.sf,opacity = 1,color = "green", group = "Simplified WDPA polygons",label = ~htmlEscape(WDPAID),weight = 1)%>%
  addPolygons(data = fishnet.sf,opacity = 1,color = "blue", group = "Fishnets",label = ~htmlEscape(poly_id),weight = 1)%>%
  addFullscreenControl() %>%
  addLayersControl(
    baseGroups = c("CartoDB", "Satellite"),
    overlayGroups = c("WDPA polygons","Simplified WDPA polygons","Fishnets"),
    options = layersControlOptions(collapsed = FALSE))

```

```{r map, message=FALSE, warning=FALSE, include=TRUE}

my_map
```

Four raster objects with varying resolutions were generated across the Bolivia region to carry out this study.
The details of the rasters thus created follows:

a) 0.1 degree (~11.1 km) = size 71.7 KB
b) 0.01 degree (~1.11 km) = size 6.9 MB
c) 0.001 degree (~111 m) = size 691.9 MB
d) 0.0005 degree (~55.5 m) = size 2.7 GB


## Results

To carry out this analysis, all the results are aggregated and group by raster resolutions. Thus, for specific raster resolution, we could recommend what could be the most suitable combination.

### 0.1 degree resolution (~11.1 km)

```{r 0.1 degree final, message=FALSE, warning=FALSE, include=FALSE}

final01 <- read.csv("data/results/03_seq_v_par_wdpa/final_01.csv")
long <- tidyr::pivot_longer(final01,
                             cols = c("WDPA_167", "Simplified_WDPA_167", "Fishnets_14338"))

result <- ggplot(data = long, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
   scale_fill_manual(values = c("#a1d99b", "#74c476", "#41ab5d", "#238b45", "#005a32",
                               "#fdbb84", "#fc8d59", "#ef6548", "#d7301f", "#990000",
                               "#9ecae1", "#6baed6", "#4292c6", "#2171b5", "#084594")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee), terra extract (te), and terra zonal (tz) using multiple cores to compute raster statistics (WDPA I)", subtitle = "(Raster spatial resolution: ~ 11 km)") +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)")

# w/o sequential
f <- final01[ ,-4]
longf <- tidyr::pivot_longer(f,
                             cols = c("WDPA_167", "Simplified_WDPA_167"))

result_new <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#a1d99b", "#74c476", "#41ab5d", "#238b45", "#005a32",
                               "#fdbb84", "#fc8d59", "#ef6548", "#d7301f", "#990000",
                               "#9ecae1", "#6baed6", "#4292c6", "#2171b5", "#084594")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee), terra extract (te), and terra zonal (tz) using multiple cores to compute raster statistics (WDPA II)", subtitle = "(Raster spatial resolution: ~ 11 km)") +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.1 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result
```

```{r 0.1 degree final plot 2, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result_new
```

### 0.01 degree resolution (~1.11 km)

```{r 0.01 degree final, message=FALSE, warning=FALSE, include=FALSE}

final001 <- read.csv("data/results/03_seq_v_par_wdpa/final_001.csv")
long <- tidyr::pivot_longer(final001,
                             cols = c("WDPA_167", "Simplified_WDPA_167", "Fishnets_14338"))

result <- ggplot(data = long, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#a1d99b", "#74c476", "#41ab5d", "#238b45", "#005a32",
                               "#fdbb84", "#fc8d59", "#ef6548", "#d7301f", "#990000",
                               "#9ecae1", "#6baed6", "#4292c6", "#2171b5", "#084594")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee), terra extract (te), and terra zonal (tz) using multiple cores to compute raster statistics (WDPA I)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)")

# w/o terra zonal (4,8,16)
f <- final001[ ,-4]
longf <- tidyr::pivot_longer(f,
                             cols = c("WDPA_167", "Simplified_WDPA_167"))

result_new <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#a1d99b", "#74c476", "#41ab5d", "#238b45", "#005a32",
                               "#fdbb84", "#fc8d59", "#ef6548", "#d7301f", "#990000",
                               "#9ecae1", "#6baed6", "#4292c6", "#2171b5", "#084594")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee), terra extract (te), and terra zonal (tz) using multiple cores to compute raster statistics (WDPA II)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.01 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result
```

_Fishnets polygon have created much spikes over the graphs, so lets remove the fishnets part and plot only WDPA and simplified WDPA polygons._

```{r 0.01 degree final plot 2, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result_new
```


### 0.001 degree resolution (~111 m)

```{r 0.001 degree final, message=FALSE, warning=FALSE, include=FALSE}

final0001 <- read.csv("data/results/03_seq_v_par_wdpa/final_0001.csv")
long <- tidyr::pivot_longer(final0001,
                             cols = c("WDPA_167", "Simplified_WDPA_167", "Fishnets_14338"))

result <- ggplot(data = long, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#a1d99b", "#74c476", "#41ab5d", "#238b45", "#005a32",
                               "#fdbb84", "#fc8d59", "#ef6548", "#d7301f", "#990000")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee) and terra extract (te) using multiple cores to compute raster statistics (WDPA)", subtitle = "(Raster spatial resolution: ~ 110 m)") +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.001 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result
```


### 0.0005 degree resolution (~55.5 m)

```{r 0.0005 degree final, message=FALSE, warning=FALSE, include=FALSE}

final00005 <- read.csv("data/results/03_seq_v_par_wdpa/final_00005.csv")
long <- tidyr::pivot_longer(final00005,
                             cols = c("WDPA_167", "Simplified_WDPA_167", "Fishnets_14338"))

result <- ggplot(data = long, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#a1d99b", "#74c476", "#41ab5d", "#238b45", "#005a32",
                               "#fdbb84", "#fc8d59", "#ef6548", "#d7301f", "#990000")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee) and terra extract (te) using multiple cores to compute raster statistics (WDPA)", subtitle = "(Raster spatial resolution: ~ 55 m)") +
  xlab("Type of Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.0005 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
result
```



## Conclusion

|**Rasters**|**size**|**Combination of functions and methods**|
|-----------|--------|----------------------------------------|
|0.1 degree (~11.1 km) |71.7 KB|1. terra zonal in single asset mode|
|0.01 degree (~1.11 km) |6.9 MB|1. terra zonal in single asset mode </br> 2. exact extract in single asset mode|
|0.001 degree (~111 m)|691.9 MB|1. for wdpa polygons, exact extract with 16 cores </br> 2. for fishnets, exact extract single asset|
|0.0005 degree (~55.5 m)|2.7 GB|1. for wdpa polygons, exact extract with 16 cores </br> 2. for fishnets, exact extract single asset|

**terra zonal did not process very small polygons in many occasions, so if your study area contains very small areas, it is recommended not to opt terra zonal.**
