---
title: "sequential vs parallelization (Fishnets)"
author: "Om Prakash Bhandari"
date: "4/13/2022"
output: workflowr::wflow_html
---

## Introduction

In this study, we are going to check how the processing time will differ - (1) sequential approach (2) parallel approach. 

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}

library(sf)
library(terra)
library(tidyr)
library(ggplot2)
library(tidyverse)
```


Four raster objects with varying resolutions were generated across the bbox extent over which the fishnets were created to carry out this study.
The details of the rasters thus created follows:

a) 0.1 degree (~11.1 km) = size 61.6 KB
b) 0.01 degree (~1.11 km) = size 6 MB
c) 0.001 degree (~111 m) = size 594.3 MB
d) 0.0005 degree (~55.5 m) = size 2.3 GB

Three different functions were used to compute these rasters over 10 different fishnet grid cells:
a) terra zonal
b) terra extract
c) exact extract

At first, the rasters were processed sequentially i.e. which uses only 1 core of the machine for the processing. Then the 
number of cores was increased to 8 and thus processing the rasters parallelly. Finally, the cores were increased up to
16. Thus, there are several results generated through this study which we will see in the bar chart below.

## **terra extract**

Here we will visualize processing results using terra extract in the form of barchart - for all resolution rasters used in study.

#### 0.1 degree resolution

```{r 0.1 degree te, message=FALSE, warning=FALSE, include=FALSE}

te01 <- read.csv("data/results/04_seq_v_par_fishnets/te_01.csv")
f <- te01[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra extract (te) using multiple cores (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 11 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- te01[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra extract (te) using multiple cores (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 11 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.1 degree te plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


#### 0.01 degree resolution

```{r 0.01 degree te, message=FALSE, warning=FALSE, include=FALSE}

te001 <- read.csv("data/results/04_seq_v_par_fishnets/te_001.csv")
f <- te001[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra extract (te) using multiple cores (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- te001[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra extract (te) using multiple cores (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.01 degree te plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


#### 0.001 degree resolution

```{r 0.001 degree te, message=FALSE, warning=FALSE, include=FALSE}

te0001 <- read.csv("data/results/04_seq_v_par_fishnets/te_0001.csv")
f <- te0001[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra extract (te) using multiple cores (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 110 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- te0001[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra extract (te) using multiple cores (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 110 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.001 degree te plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


#### 0.0005 degree resolution

```{r 0.0005 degree te, message=FALSE, warning=FALSE, include=FALSE}

te00005 <- read.csv("data/results/04_seq_v_par_fishnets/te_00005.csv")
f <- te00005[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra extract (te) using multiple cores (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 55 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- te00005[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra extract (te) using multiple cores (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 55 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.0005 degree te plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```

### Findings

**1. for coarse resolution raster, sequential method performs better**  
**2. with increase in resolution, parallelization seems to perform way better**  
**3. however, usage of 8 cores and 16 cores did not affect the result significantly - so it is better to process with 8 cores**  

## **exact extract**

Here we will visualize processing results in barchart - for all resolution rasters used in study 

#### 0.1 degree resolution

```{r 0.1 degree ee, message=FALSE, warning=FALSE, include=FALSE}

ee01 <- read.csv("data/results/04_seq_v_par_fishnets/ee_01.csv")
f <- ee01[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by exact extract (ee) using multiple cores (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 11 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- ee01[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by exact extract (ee) using multiple cores (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 11 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.1 degree ee plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


#### 0.01 degree resolution

```{r 0.01 degree ee, message=FALSE, warning=FALSE, include=FALSE}

ee001 <- read.csv("data/results/04_seq_v_par_fishnets/ee_001.csv")
f <- ee001[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by exact extract (ee) using multiple cores (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- ee001[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by exact extract (ee) using multiple cores (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.01 degree ee plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


#### 0.001 degree resolution

```{r 0.001 degree ee, message=FALSE, warning=FALSE, include=FALSE}

ee0001 <- read.csv("data/results/04_seq_v_par_fishnets/ee_0001.csv")
f <- ee0001[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by exact extract (ee) using multiple cores (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 110 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- ee0001[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by exact extract (ee) using multiple cores (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 110 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.001 degree ee plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


#### 0.0005 degree resolution

```{r 0.0005 ee degree, message=FALSE, warning=FALSE, include=FALSE}

ee00005 <- read.csv("data/results/04_seq_v_par_fishnets/ee_00005.csv")
f <- ee00005[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by exact extract (ee) using multiple cores (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 55 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- ee00005[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by exact extract (ee) using multiple cores (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 55 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.0005 degree ee plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


### Findings

**1. for coarse resolution raster, sequential method performs better**  
**2. for higher resolution raster too, sequential method performs bettter**  
**3. the only case where parallelization (particularly 16 cores) slightly outperforms sequential method is - when the study area polygon is excessively large (~5000 or more sqkm) and rasters are of very high resolution**


## **terra zonal**

Here we will visualize processing results in barchart - for all resolution rasters used in study 

#### 0.1 degree resolution

```{r 0.1 degree tz, message=FALSE, warning=FALSE, include=FALSE}

tz01 <- read.csv("data/results/04_seq_v_par_fishnets/tz_01.csv")
f <- tz01[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra zonal (tz) using multiple cores (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 11 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- tz01[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra zonal (tz) using multiple cores (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 11 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.1 degree tz plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


#### 0.01 degree resolution

```{r 0.01 degree tz, message=FALSE, warning=FALSE, include=FALSE}

tz001 <- read.csv("data/results/04_seq_v_par_fishnets/tz_001.csv")
f <- tz001[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra zonal (tz) using multiple cores (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- tz001[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken to compute raster statistics by terra zonal (tz) using multiple cores (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.01 degree tz plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


### Findings

**1. for coarse resolution rasters (~1 km) and increase in number of polygons - terra zonal performs better in sequential mode**  
**2. processing extremely big datasets result in poor performance sometimes leading to failure of R session - so it is not recommended to process very high resolution ratsers using terra zonal**


## summarise

|functions|findings|
|---------|--------|
|terra extract|1. for coarse resolution raster, sequential method performs better </br> 2. with increase in resolution, parallelization seems to perform way better </br> 3. however, usage of 8 cores and 16 cores did not affect the result significantly - so it is better to process with 8 cores|
|exact extract|1. for coarse resolution raster, sequential method performs better </br> 2. for higher resolution raster too, sequential method performs bettter </br> 3. the only case where parallelization (particularly 16 cores) slightly outperforms sequential method is - when the study area polygon is excessively large (~5000 or more sqkm) and rasters are of very high resolution|
|terra zonal|1. for coarse resolution rasters (~1 km) and increase in number of polygons - terra zonal performs better in sequential mode </br> 2. processing extremely big datasets result in poor performance sometimes leading to failure of R session - so it is not recommended to process very high resolution ratsers using terra zonal|


## **most suitable combination?**

We only saw above how the three functions perform over the varying raster resolutions and varying study area polygons. So, now is the time to make recommendations for the ideal combination of functions and methods (sequential or parallelization). To carry out this analysis, all the results are aggregated and group by raster resolutions. Thus, for specific raster resolution, we could recommend what could be the most suitable combination.

### 0.1 degree resolution (~11.1 km)

```{r 0.1 degree final, message=FALSE, warning=FALSE, include=FALSE}

final01 <- read.csv("data/results/04_seq_v_par_fishnets/final_01.csv")
f <- final01[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#74c476", "#41ab5d", "#238b45",
                               "#fc8d59", "#ef6548", "#d7301f",
                               "#6baed6", "#4292c6", "#2171b5")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee), terra extract (te), and terra zonal (tz) using multiple cores to compute raster statistics (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 11 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- final01[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#74c476", "#41ab5d", "#238b45",
                               "#fc8d59", "#ef6548", "#d7301f",
                               "#6baed6", "#4292c6", "#2171b5")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee), terra extract (te), and terra zonal (tz) using multiple cores to compute raster statistics (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 11 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.1 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


#### the best combination

**1. terra zonal in sequential mode**


### 0.01 degree resolution (~1.11 km)

```{r 0.01 degree final, message=FALSE, warning=FALSE, include=FALSE}

final001 <- read.csv("data/results/04_seq_v_par_fishnets/final_001.csv")
f <- final001[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#74c476", "#41ab5d", "#238b45",
                             "#fc8d59", "#ef6548", "#d7301f",
                             "#6baed6", "#4292c6", "#2171b5")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee), terra extract (te) and terra zonal (tz) using multiple cores to compute raster statistics (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- final001[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#74c476", "#41ab5d", "#238b45",
                             "#fc8d59", "#ef6548", "#d7301f",
                             "#6baed6", "#4292c6", "#2171b5")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee), terra extract (te) and terra zonal (tz) using multiple cores to compute raster statistics (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")

# w/o terra zonal
final001_new <- final001[-c(2:3), ]
f <- final001_new[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf_new <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#74c476", "#41ab5d", "#238b45",
                             "#fc8d59", "#ef6548", "#d7301f",
                             "#4292c6")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee), terra extract (te) and terra zonal (tz) using multiple cores to compute raster statistics (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- final001_new[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results_new <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#74c476", "#41ab5d", "#238b45",
                             "#fc8d59", "#ef6548", "#d7301f",
                             "#4292c6")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee), terra extract (te) and terra zonal (tz) using multiple cores to compute raster statistics (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 1 km)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.01 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


Since, we couldn't visualize properly the lower range of values, lets remove terra zonal 8 and 16 cores from the graph.

```{r 0.01 degree new final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf_new
results_new
```


#### the best combination

**1. for large number of polygons and small areas upto ~100 sqkm, terra extract in parallelization mode (16 cores)**  
**2. for small number of polygons and large areas >100 sqkm, terra zonal in sequential mode**


### 0.001 degree resolution (~111 m)

```{r 0.001 degree final, message=FALSE, warning=FALSE, include=FALSE}

final0001 <- read.csv("data/results/04_seq_v_par_fishnets/final_0001.csv")
f <- final0001[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#41ab5d", "#238b45", "#005a32",
                               "#ef6548", "#d7301f", "#990000")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee) and terra extract (te) using multiple cores to compute raster statistics (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 110 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- final0001[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#41ab5d", "#238b45", "#005a32",
                               "#ef6548", "#d7301f", "#990000")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee) and terra extract (te) using multiple cores to compute raster statistics (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 110 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")

# w/o terra extract (sequential)
final0001_new <- final0001[2:6, ]
f <- final0001_new[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf_new <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#74c476", "#41ab5d", "#238b45",
                             "#fc8d59", "#ef6548")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee) and terra extract (te) using multiple cores to compute raster statistics (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 110 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- final0001_new[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results_new <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#74c476", "#41ab5d", "#238b45",
                             "#fc8d59", "#ef6548")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee) and terra extract (te) using multiple cores to compute raster statistics (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 110 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.001 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```

Since, terra extract in sequential mode has very high values compared to other combination of functions and cores, we will not take it into consideration and visualize only the close values.

```{r 0.001 degree new final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf_new
results_new
```


#### the best combination

**1. for large number of polygons and areas upto ~1000 sqkm, exact extract in sequential mode**  
**2. for extremely large areas >1000 sqkm, exact extract in parallelization mode (16 cores), however the difference with sequential is not so significant (~1 secs)**


### 0.0005 degree resolution (~55.5 m)

```{r 0.0005 degree final, message=FALSE, warning=FALSE, include=FALSE}

final00005 <- read.csv("data/results/04_seq_v_par_fishnets/final_00005.csv")
f <- final00005[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#41ab5d", "#238b45", "#005a32",
                               "#ef6548", "#d7301f", "#990000")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee) and terra extract (te) using multiple cores to compute raster statistics (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 55 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- final00005[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#41ab5d", "#238b45", "#005a32",
                               "#ef6548", "#d7301f", "#990000")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee) and terra extract (te) using multiple cores to compute raster statistics (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 55 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")

# w/o terra extract (sequential)
final00005_new <- final00005[2:6, ]
f <- final00005_new[,1:6]
longf <- tidyr::pivot_longer(f,
                             cols = c("grids_255020", "grids_127600", "grids_51336", "grids_25839", "grids_12950"))
resultf_new <- ggplot(data = longf, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#74c476", "#41ab5d", "#238b45",
                             "#fc8d59", "#ef6548")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee) and terra extract (te) using multiple cores to compute raster statistics (Fishnets I)", subtitle = "(Raster spatial resolution: ~ 55 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")


s <- final00005_new[,c(1,7:11)]
longs <- tidyr::pivot_longer(s,
                             cols = c("grids_6600", "grids_2688", "grids_1350", "grids_294", "grids_157"))
results_new <- ggplot(data = longs, aes(x = reorder(name, -value), y = value, fill = Cores)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#74c476", "#41ab5d", "#238b45",
                             "#fc8d59", "#ef6548")) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(), plot.title = element_text(size = rel(1.25), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.1), hjust = 0.5)) +
  labs(title = "Time taken by exact extract (ee) and terra extract (te) using multiple cores to compute raster statistics (Fishnets II)", subtitle = "(Raster spatial resolution: ~ 55 m)") +
  xlab("Polygons") +
  ylab("Time elapsed (s)")
```

```{r 0.0005 degree final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf
results
```


Since, terra extract in sequential mode has very high values compared to other combination of functions and cores, we will not take it into consideration and visualize only the close values.

```{r 0.0005 degree new final plot, message=FALSE, warning=FALSE, include=TRUE, fig.width=12}
resultf_new
results_new
```


#### the best combination

**1. for large number of polygons and areas upto ~1000 sqkm, exact extract in sequential mode**  
**2. for extremely large areas >1000 sqkm, exact extract in parallelization mode (16 cores)**


## summarise

|**Rasters**|**size**|**Combination of functions and methods**|
|-----------|--------|----------------------------------------|
|0.1 degree (~11.1 km) |61.6 KB|1. terra zonal in single asset mode|
|0.01 degree (~1.11 km) |6 MB|1. for large number of polygons and small areas upto ~100 sqkm, terra extract in parallelization mode (16 cores) </br> 2. for small number of polygons and large areas >100 sqkm, terra zonal in single asset mode|
|0.001 degree (~111 m)|594.3 MB|1. for large number of polygons and areas upto ~1000 sqkm, exact extract in single asset mode </br> 2. for extremely large areas >1000 sqkm, exact extract in parallelization mode (16 cores), however the difference with single asset mode is not so significant (~1 secs)|
|0.0005 degree (~55.5 m)|2.3 GB|1. for large number of polygons and areas upto ~1000 sqkm, exact extract in single asset mode </br> 2. for extremely large areas >1000 sqkm, exact extract in parallelization mode (16 cores)|

